Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-01-31 manuscript/HIAs.qmd
validate fail

@@

>>>>

# Calulculate area
infra <- infra |>
  mutate(area = geometry |> st_area(),
         area_km = area |> set_units("km2"))

# show the summered area per HIA
infra |>
  as_tibble() |>
  group_by(infrastructureIndex) |>
  summarise(area_km = sum(area_km)) |>
  ggplot(aes(x = infrastructureIndex,
             y = area_km)) +
  geom_col()

@@

```{r HIA-validate}
#| eval: false

# Now I want to see if the indicator values depend on the HIA is a predictable
# way to justify the stratification
corrCheck <- st_intersection(naturetypes, infra)

corrCheck |>
  as_tibble() |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  count(i_ADSV_fct, infrastructureIndex) |>
  View()

# A first look
corrCheck |>
  mutate(i_ADSV_fct = floor(round(i_ADSV*10,2))/10,
         i_alien_fct = floor(round(i_alien*10,2))/10) |>
  pivot_longer(cols = c(i_ADSV_fct, i_alien_fct),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = factor(round(indicatorValue,2))))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_fill_brewer(palette = "RdYlGn")+
  facet_grid(indicator~year)
```

```{r}
#| eval: false

# Using a colour gradient emphasises the first colour (dark green).
# Lets try discrete colours, and merge som classes to simplify
corrCheck |>
  pivot_longer(cols = c(i_ADSV, i_alien),
               values_to = "indicatorValue",
               names_to = "indicator") |>
  filter(!is.na(indicatorValue)) |>
  mutate(condition = case_when(
    indicatorValue <0.6 ~"Poor",
    indicatorValue <0.8 ~"Intermediate",
    .default = "Good"
  )) |>
  ggplot(aes(
    x = factor(infrastructureIndex), 
    fill = condition))+
  geom_bar(
    position="fill"
    )+
  theme_bw(base_size = 12)+
  guides(fill = guide_legend("Scaled indicator values"))+
  ylab("Fraction of data points")+
  xlab("HIA")+
  scale_y_log10() +
  scale_fill_viridis_d(alpha = .8) +
  facet_grid(~indicator)
